---
title: "R Package: scorelingam"
output: 
  md_document:
    toc: true
---

# Description

This package implements the estimation of a topological ordering for a Linear Structural Equation Model (SEM) with non-Gaussian errors, as outlined in [Ruiz et. al (2022+)](https://openreview.net/forum?id=4pCjIGIjrt):


> G Ruiz, OH Madrid-Padilla, Q Zhou. "Sequentially learning the topological ordering of directed acyclic graphs with likelihood ratio scores." Transactions of Machine Learning Research, 2022+. [[Link](https://openreview.net/forum?id=4pCjIGIjrt)]

See below for some examples. 

# Install package from Github. 

```{r, echo = T, eval=F}
# install devtools, if needed. 
require(devtools)
# install the scorelingam package. 
devtools::install_github(repo='gabriel-ruiz/scorelingam')
```
```{r}
# load scorelingam once installed
library(scorelingam)
```

# An example for our sorting procedure: $p=5$


## Generate data 
### Linear SEM Parameters
```{r}
p = 5;numRoots = 1; numParentsMin = 1; numParentsMax = 2
scaleParam = runif(n=p,min=0.5,max=1.2)
causalOrder = 5:1
lingamParams = randomWeightedAdjacencyMatrix(p=p,num.roots=numRoots,pa.min=numParentsMin,pa.max=numParentsMax,
                    pa.wt.min = 0.25,pa.wt.max = 0.9,prob.pos = 0.5,perm = causalOrder)
# weighted adjacency matrix
print(lingamParams$B)
```
### Data matrix with $n=5000$
When generating the data matrix, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r}
n = 5000
X = generateData(B=lingamParams$B,scale = scaleParam,perm = causalOrder,n = n,family = 'laplace')
dim(X)
```

## Obtain an estimated ordering and check proportion of parents sorted after a child
When estimating a topological ordering for the linear SEM, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r order-step-small-graph}
# neighborhoods specified to be all other nodes
mbhat = lapply(1:ncol(X),function(j){(1:ncol(X))[-j]}) 
estOrder = scorelingam(X=X,mb=mbhat,numUpdates=ncol(X),family='laplace')
# check errors (ideally close to zero)
checkSortingErrors(estOrder=estOrder,A=lingamParams$B)
```

## Check accuracy of estimated weighted adjacency matrix

```{r}
paHat = getParents(mb=mbhat,ordering=estOrder)
(Bhat = getWeights(X=scale(X,scale=F,center=T),pa=paHat))
# maximum entry-wise difference in absolute value
norm(x=lingamParams$B-Bhat,type = 'i')
```


# A higher dimensional example: $p=10,000$
## Generate data 
### Linear SEM Parameters
```{r,cache=F}
p = 10000;numRoots = 100; numParentsMin = 1; numParentsMax = 2
scaleParam = runif(n=p,min=0.5,max=1.2)
causalOrder = c(seq(2,p,by=2),seq(1,p-1,by=2))
lingamParams = randomWeightedAdjacencyMatrix(p=p,num.roots=numRoots,pa.min=numParentsMin,pa.max=numParentsMax,
                    pa.wt.min = 0.25,pa.wt.max = 0.9,prob.pos = 0.5,perm = causalOrder)
```
### Data matrix with $n=5000$
When generating the data matrix, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r,cache=F}
n = 5000
X = generateData(B=lingamParams$B,scale = scaleParam,perm = causalOrder,n = n,family = 'laplace')
dim(X)
```

## Obtain an estimated ordering and time the algorithm
When estimating a topological ordering for the Linear SEM, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r,cache=F}
# neighborhoods specified to be true markov blankets
mbhat = moralize(lingamParams$B) 
#
start = Sys.time()
estOrder = scorelingam(X=X,mb=mbhat,numUpdates=10,family='laplace')
end = Sys.time()
difftime(end,start,units='mins')

```

## Check accuracy of estimated ordering

```{r,cache=F}
# check sorting errors (ideally close to zero)
# proportion of parents sorted after child
checkSortingErrors(estOrder=estOrder,A=lingamParams$B)
```

