---
title: "Some Examples of our Sorting Procedure"
output: 
  pdf_document:
    toc: true
---

# An example for our parametric LiNGAM sorting procedure ($p=5$)

## Compiling necessary code for this example. 

The below code chunk assumes that the helper function \texttt{R} file and algorithm source code are in the same directory as this \texttt{R} markdown file. 
```{r}
# helper functions for generating LiNGAM data and checking correctness of order
source("source/scorelingam/helperFunctions.r") 
require(Rcpp)
# compile sorting algorithm source, and allow it to be used in R
Rcpp::sourceCpp("source/scorelingam/source.cpp") 
```

## Generate data from a LiNGAM
### LiNGAM Parameters
```{r}
p <- 5;numRoots <- 1; numParentsMin <- 1; numParentsMax <- 2
scaleParam <- runif(n=p,min=0.5,max=1.2)
causalOrder <- 5:1
lingamParams <- rand.wtd.adj.mat(p=p,num.roots=numRoots,pa.min=numParentsMin,pa.max=numParentsMax,
                    pa.wt.min = 0.25,pa.wt.max = 0.9,prob.pos = 0.5,perm = causalOrder)
# weighted adjacency matrix
print(lingamParams$B)
```
### Data matrix with $n=5000$
When generating the data matrix, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r}
n <- 5000
X <- genSCM.data(B=lingamParams$B,shape = scaleParam,perm = causalOrder,n = n,family = 'laplace')
dim(X)
```

## Obtain an estimated ordering and check proportion of parents sorted after a child
When estimating a topological ordering for the LiNGAM, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r}
# neighborhoods specified to be all other nodes
mbhat <- lapply(1:ncol(X),function(j){(1:ncol(X))[-j]}) 
(estOrder <- sort_llrmbCPP(Xmat=X,mb=mbhat,numUpdates=ncol(X),family='laplace'))
# check errors (ideally close to zero)
check.valid.sort(estOrder=estOrder,M=t(lingamParams$B))
```

## Check accuracy of estimated weighted adjacency matrix

```{r}
(Bhat <- getParams(X=X,mbhat=mbhat,ordering=estOrder)$Bhat)
# maximum entry-wise difference in absolute value
norm(x=lingamParams$B-Bhat,type = 'i')
```


# A higher dimensional example ($p=10,000$)
## Generate data from a LiNGAM 
### LiNGAM Parameters
```{r,cache=F}
p <- 10000;numRoots <- 100; numParentsMin <- 1; numParentsMax <- 2
scaleParam <- runif(n=p,min=0.5,max=1.2)
causalOrder <- c(seq(2,p,by=2),seq(1,p-1,by=2))
lingamParams <- rand.wtd.adj.mat(p=p,num.roots=numRoots,pa.min=numParentsMin,pa.max=numParentsMax,
                    pa.wt.min = 0.25,pa.wt.max = 0.9,prob.pos = 0.5,perm = causalOrder)
```
### Data matrix with $n=5000$
When generating the data matrix, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r,cache=F}
n <- 5000
X <- genSCM.data(B=lingamParams$B,shape = scaleParam,perm = causalOrder,n = n,family = 'laplace')
dim(X)
```

## Obtain an estimated ordering and time the algorithm
When estimating a topological ordering for the LiNGAM, the possible options for the family argument are 'laplace', 'logistic', and 't', in which case the additional argument df is needed (df=10 is the default)
```{r,cache=F}
# neighborhoods specified to be true markov blankets
mbhat <- moralize(lingamParams$B) 
#
start <- Sys.time()
estOrder <- sort_llrmbCPP(Xmat=X,mb=mbhat,numUpdates=10,family='laplace')
end <- Sys.time()
difftime(end,start,units='mins')

```

## Check accuracy of estimated ordering

```{r,cache=F}
# check sorting errors (ideally close to zero)
# proportion of parents sorted after child
check.valid.sort(estOrder=estOrder,M=t(lingamParams$B))
```


```